---
title: "music_company Project"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

What Makes a Good Playlist?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(scipen=999)

install.packages("readr")
library(readr)

install.packages("tidyr")
library(tidyr)

install.packages("ggplot2")
library(ggplot2)

install.packages("dplyr")
library(dplyr)

install.packages("cluster")
library(cluster)

install.packages("factoextra")
library(factoextra)

install.packages("purrr")
library(purrr)

install.packages("plotly")
library(plotly)

install.packages("psych")
library(psych)

install.packages("MASS")
library(MASS)


#install.packages("stats")
#libnrary(stats)

playlist_data <- read_delim("Downloads/playlist_summary_external.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)

#rename the columns to be more intuitive
colnames(playlist_data)[5:7]<- c("daily_active_users", "weekly_active_users", "monthly_active_users")
View(playlist_data)

nrow(playlist_data)
ncol(playlist_data)

###Create a second table that separates the tokens, including the playlist url, the playlist owner, streams, daily active users, and then
#the tokens split up into different columns

#Only going up to 10 tokens as a max
token_df<- separate(playlist_data, 'tokens', paste("tokens", 1:5, sep="_"), sep=",", extra="drop")
View(token_df)



```

Questions to Answer about the Data Set Through EDA
1. What genre/ genre combinations are the most common?
2. What genre/ genre combinations are the best by streams? 
3. What genre/ genre combinations are the best by daily active users?
4. Are any of the variables highly correlated?
5. Is there a relationship with the number of tracks to the streams and daily active users.


Exploratory Data Analysis

1. Distribution Charts
2. Percentiles
3. Frequency Tables for the Genres and Moods.

```{r}


#Distribution of Numeric Variables

outliers<- playlist_data[c(152033, 163727),]


summary(outliers)


###############3

playlist_data<- playlist_data[-c(152033, 163727),]

playlist_no_outliers<- playlist_data

playlist_data<- subset(playlist_data, playlist_data$streams != 0)

#todays streams 
quantile_list<- c(.10, .20, .30, .40, .50, .60, .70, .80, .90, 1)
top_5_quantile_list<- c(.90, .95, .99, 1)
quantile(playlist_data$streams, quantile_list)

#todays streams over 30 seconds
quantile(playlist_data$stream30s, quantile_list)
quantile(playlist_data$stream30s, top_5_quantile_list)

nrow(subset(playlist_data, playlist_data$stream30s >= 210))


ggplot(playlist_data, aes(x = stream30s, y = daily_active_users)) + 
  geom_point(shape=23, fill="blue", size=3) +
  ggtitle("Streams by Daily Active Users") +
  xlab("Streams Over 30 Seconds") +
  ylab("Daily Active Users") +
  geom_vline(aes(xintercept = 54))
  
  
# daily active users
quantile(playlist_data$daily_active_users, top_5_quantile_list)
# weekly active users
quantile(playlist_data$weekly_active_users, top_5_quantile_list)
#monthly_active_users
quantile(playlist_data$monthly_active_users, top_5_quantile_list)
# monthly active user previous month
quantile(playlist_data$mau_previous_month, top_5_quantile_list)
#monthly active users both months
quantile(playlist_data$mau_both_months, top_5_quantile_list)
#users
quantile(playlist_data$users, top_5_quantile_list)
#skippers
quantile(playlist_data$skippers, top_5_quantile_list)
#n_tracks
quantile(playlist_data$n_tracks, top_5_quantile_list)
#n_local_tracks
quantile(playlist_data$n_local_tracks, top_5_quantile_list)
#n_artists
quantile(playlist_data$n_artists, top_5_quantile_list)
#albums
quantile(playlist_data$n_albums, top_5_quantile_list)
#monthly streams 30 seconds
quantile(playlist_data$monthly_stream30s, top_5_quantile_list)
#monthly owner streams
quantile(playlist_data$monthly_owner_stream30s, top_5_quantile_list)


#Correlation Plot
cor(playlist_data[3:11, 13:18])

playlist_data$long_stream_perc<- playlist_data$stream30s/playlist_data$streams
playlist_data$user_long_stream_ratio<- playlist_data$stream30s/playlist_data$daily_active_users
playlist_data$mom_change<- (playlist_data$mau_previous_month- playlist_data$monthly_active_users)/playlist_data$mau_previous_month

playlist_data$mom_change<- ifelse(playlist_data$mom_change == "-Inf", 0 , playlist_data$mom_change)

playlist_data$user_long_stream_ratio<- ifelse(playlist_data$user_long_stream_ratio == "NaN", 0, playlist_data$user_long_stream_ratio)

corr_df<- as.data.frame(cor(playlist_data[,c(3:11, 13:18,26:28)]))
#corr_df<- subset(corr_df, corr_df == 1.00000000)

corr_long_stream<- as.data.frame(cbind(row.names(corr_df), corr_df$long_stream_perc))
colnames(corr_long_stream)<- c("Factor", "Correlation Level")
corr_long_stream<- subset(corr_long_stream, corr_long_stream$`Correlation Level` != 1)


ggplot(corr_long_stream, aes(x = Factor, y = sort(`Correlation Level`, decreasing = TRUE))) +
  geom_bar(stat = "identity") +
  scale_y_continuous(name = "Correlation Value", limits = c(-.2, .2))


corr_user_ratio<- as.data.frame(cbind(row.names(corr_df), corr_df$user_long_stream_ratio))
colnames(corr_user_ratio)<- c("Factor", "Correlation Level")
corr_user_ratio<- subset(corr_user_ratio, corr_user_ratio$`Correlation Level` != 1)

corr_mom_change<- as.data.frame(cbind(row.names(corr_df), corr_df$mom_change))
colnames(corr_mom_change)<- c("Factor", "Correlation Level")
corr_mom_change<- subset(corr_mom_change, corr_mom_change$`Correlation Level` != 1)


ggplot(corr_df[], aes(x = row.names(corr_df), y = sort(long_stream_perc, decreasing = TRUE))) +
  geom_bar(stat = "identity")


#Create Summary Tables by Genre 1

genre_stats_1<- playlist_data %>%
group_by(genre_1) %>%
summarise(mean_streams = mean(streams), mean_stream30s = mean(stream30s), mean_tracks = mean(n_tracks), mean_local_tracks = mean(n_local_tracks),
          mean_artists = mean(n_local_tracks), mean_albums = mean(n_albums), mean_monthly_streams = mean(monthly_stream30s), 
          mean_owner_stream_30s = mean(monthly_owner_stream30s))

#Genre 2 

genre_stats_2<- playlist_data %>%
group_by(genre_2) %>%
summarise(mean_streams = mean(streams), mean_stream30s = mean(stream30s), mean_tracks = mean(n_tracks), mean_local_tracks = mean(n_local_tracks),
          mean_artists = mean(n_local_tracks), mean_albums = mean(n_albums), mean_monthly_streams = mean(monthly_stream30s), 
          mean_owner_stream_30s = mean(monthly_owner_stream30s))


#Genre 3

genre_stats_3<- playlist_data %>%
group_by(genre_3) %>%
summarise(mean_streams = mean(streams), mean_stream30s = mean(stream30s), mean_tracks = mean(n_tracks), mean_local_tracks = mean(n_local_tracks),
          mean_artists = mean(n_local_tracks), mean_albums = mean(n_albums), mean_monthly_streams = mean(monthly_stream30s), 
          mean_owner_stream_30s = mean(monthly_owner_stream30s))


#Mood 1

mood_stats_1<- playlist_data %>%
group_by(mood_1) %>%
summarise(mean_streams = mean(streams), mean_stream30s = mean(stream30s), mean_tracks = mean(n_tracks), mean_local_tracks = mean(n_local_tracks),
          mean_artists = mean(n_local_tracks), mean_albums = mean(n_albums), mean_monthly_streams = mean(monthly_stream30s), 
          mean_owner_stream_30s = mean(monthly_owner_stream30s))

#Mood 2

mood_stats_2<- playlist_data %>%
group_by(mood_1) %>%
summarise(mean_streams = mean(streams), mean_stream30s = mean(stream30s), mean_tracks = mean(n_tracks), mean_local_tracks = mean(n_local_tracks),
          mean_artists = mean(n_local_tracks), mean_albums = mean(n_albums), mean_monthly_streams = mean(monthly_stream30s), 
          mean_owner_stream_30s = mean(monthly_owner_stream30s))


#Mood 3

mood_stats_3<- playlist_data %>%
group_by(mood_3) %>%
summarise(mean_streams = mean(streams), mean_stream30s = mean(stream30s), mean_tracks = mean(n_tracks), mean_local_tracks = mean(n_local_tracks),
          mean_artists = mean(n_local_tracks), mean_albums = mean(n_albums), mean_monthly_streams = mean(monthly_stream30s), 
          mean_owner_stream_30s = mean(monthly_owner_stream30s))



#histogrms of the metrics, greater than 0 
hist(which(playlist_data$streams > 0))
hist(which(playlist_data$stream30s > 0))
hist(which(playlist_data$daily_active_users > 0))
hist(which(playlist_data$weekly_active_users > 0))
hist(which(playlist_data$monthly_active_users > 0))

########


monthly_funnel <- plot_ly() 
monthly_funnel <- monthly_funnel %>%
  add_trace(
  type = "funnel",
  y = c("Monthly Active Users","Weekly Active Users", "Daily Active Users"),
  x = c(mean(outliers$monthly_active_users), 
        mean(outliers$weekly_active_users),
        mean(outliers$daily_active_users)
        )) 
monthly_funnel <- monthly_funnel %>%
  layout(yaxis = list(categoryarray = c("Monthly Active Users","Weekly Active Users", "Daily Active Users")))
monthly_funnel



#######

View(outliers)


########

ggplot(playlist_data, aes(x = streams, y = daily_active_users)) + 
  geom_point() +
  ggTitle("Streams by Daily Active Users") +
  ylab


ggplot(playlist_data, aes(x = stream30s, y = daily_active_users)) + 
  geom_point(shape=23, fill="blue", size=3) +
  ggtitle("Streams by Daily Active Users") +
  xlab("Streams Over 30 Seconds") +
  ylab("Daily Active Users")


```

The question what makes a good playlist presents an unsupervised problem since there is not singular succinct metric (as defined in the data) that makes a great playlist. Because of this, the answer to this question will be approached using an unsupervised clustering method, as well as a supervised method.


Unsupervised Analysis
1. K- Means Clustering Algorithm

```{r}

playlist_data<- playlist_no_outliers
set.seed(227)

# function to compute total within-cluster sum of square 
within_ss_function <- function(k) {
  kmeans(playlist_data[c(3:11,13:18)], k, nstart = 10)$tot.withinss
}

num_clusters <- 2:15

within_ss_values<- map_dbl(num_clusters, within_ss_function)

plot(num_clusters, within_ss_values,
       type="b", pch = 19, frame = FALSE, 
       xlab="Number of Clusters",
       ylab="Total Within-Clusters Sum of Squares")

playlist_clusters<- kmeans(playlist_data[c(3:11,13:18)], 5)
table(playlist_clusters$cluster)

#cluster_summary<- playlist_clusters %>% split(.$cluster) %>% map(summary)


playlist_data$long_stream_perc<- playlist_data$stream30s/playlist_data$streams
playlist_data$user_long_stream_ratio<- playlist_data$stream30s/playlist_data$daily_active_users
playlist_data$mom_change<- (playlist_data$mau_previous_month- playlist_data$monthly_active_users)/playlist_data$mau_previous_month

playlist_data$mom_change<- ifelse(playlist_data$mom_change == "-Inf", 0 , playlist_data$mom_change)

playlist_data$user_long_stream_ratio<- ifelse(playlist_data$user_long_stream_ratio == "NaN", 0, playlist_data$user_long_stream_ratio)


cluster_df<- cbind(playlist_clusters$cluster, playlist_data)

describeBy(cluster_df[,-c(1:3)], cluster_df[1])
#cluster_summary


#Genre Majority
table(as.data.frame(subset(cluster_df, cluster_df$`playlist_clusters$cluster` == 1)[,21]))
table(as.data.frame(subset(cluster_df, cluster_df$`playlist_clusters$cluster` == 2)[,21]))
table(as.data.frame(subset(cluster_df, cluster_df$`playlist_clusters$cluster` == 3)[,21]))
table(as.data.frame(subset(cluster_df, cluster_df$`playlist_clusters$cluster` == 4)[,21]))
table(as.data.frame(subset(cluster_df, cluster_df$`playlist_clusters$cluster` == 5)[,21]))


#Mood Majority
table(as.data.frame(subset(cluster_df, cluster_df$`playlist_clusters$cluster` == 1)[,24]))
table(as.data.frame(subset(cluster_df, cluster_df$`playlist_clusters$cluster` == 2)[,24]))
table(as.data.frame(subset(cluster_df, cluster_df$`playlist_clusters$cluster` == 3)[,24]))
table(as.data.frame(subset(cluster_df, cluster_df$`playlist_clusters$cluster` == 4)[,24]))
table(as.data.frame(subset(cluster_df, cluster_df$`playlist_clusters$cluster` == 5)[,24]))



```

Notes:

Cluster 1: Long Term Pop Success

This playlist is in it for the long run. The month over month change in the monthly active users on the playlist is +8%, which is a low increase comparatively to the other clusters, meaning that the growth is stagnent, but the dedication of the users is strong.
Although this cluster only has 13 playlists, these are successful playlists that have a lot of longevity and will be successful for a long term. This segment is made primarily made up of Excited Pop and Indie Rock playlists, and over 50% of the of the streams are for longer than 30 seconds. This segment includes the highest volume of streams, users, long streams, and the lowest month over month change. The most successful playlists are in this segment.

13 Playlists 
Excited, Pop and Indie Rock Majority
Over 50 Long Stream Percentage
About 10 songs Per User
~141 Tracks
greater than 350,000 monthly streams
greater than 20,000 daily users
12,000 skippers


Cluster 2: 
Personal Playlists Segments- 
This segment includes the highest number of playlists at over 40,000. This segment is made up of majority of personal playlists. There is no genre or mood majority. There is also very little coverage of this playlist outside of the owner, since the owner makes up a majority of the listens in this playlist (about 33% of streams come from the owner). This allows metrics such as the percentage of long streams to be very high, while the average number of listeners and the average number of songs is very low (on average 1 daily active user).

403122
Genre Fluid/ No Genre Majority
Personal Playlists- these are playlists that are made for the listening enjoyment of a particular person or close group- High Owner Streams
There is very little coverage of this playlist outside of the owner of the playlist. This allows metrics such as the long
stream percentage to be relatively high at 65%, but is mitigated due to the low volume of users and streams (15 users monthly, 10 steams more than 30 seconds per day)
Average Streams Less than 20
Daily Active Users ~1
~40 users
~33% of monthly streams come from owners
~Monthly Active Users less than 5 for both months
High Owner Streams


Cluster 3: High Popularity, Mid Range Longevity. 

This cluster contains 32 playlists, another low volume segment, but is also the segment with the seconds highest number of streams, users, and monthly active users. The largest difference between this segment and the larger segment is that the average month over change for this segment is increasing at a higher rate than the segment with the most popular playlists. A 25% month over month increase in over 30 second streams. This playlist is majority pop, Indie rock, and dance with an overall excited mood. Good user retention as users are on average listening to 9 songs in this playlist. No artist also appears in the playlist more than once. This plylist has a little less than half of the daily/weekly/monthly streams of the high rollers.

32 Playlists
Excited Pop, Indie Rock, and Dance Majority- High Tier Success
greater than 150,000 daily streams
about 10,000 daily active users
each user is listerning to abotu 9 songs on this playlist
the owner does not make up a majority of the streams.
High artist range, as no artists has more than 1 song on the playlist.
Monthly streams greater than 2 million
High long stream percentage
A little less than half of the daily/weekly/monthly streams of the first high rollers cluster


Cluster 4:

Defiant Low- Tier Playlist
This segment contains playlists that can be considered low-tier. They do not have the daily/weekly/monthly volume of the playlists in the high roller segments, but they still pull volumes that are in the top 2% of the playlists. There are 151 playlists in this range. Similarly to the secondary high rollers segment, this cluster also has a low average month over month change in streams. There is also high long stream percentage and on average users are listening to about 11 songs. The mood of the playlists in this segment is primarily defiant and the genres are primarily Indie Rock and Pop. The playlist in this cluster also have a higher range for key metrics such as streams and over 30 seconds streams compared to the other clusters. Although the volumes differ, there are patterns in terms of retention that make the playlists in this cluster more similar to each other than those in other playlists.


151 Playlists
Defiant Indie Rock, Pop Majority
Average Success 


Cluster 5
46 Playlists
Excited Indie Rock and Pop
Mid-Tier High Growth

This segment is mid-tier are it falls right between the high tier and the low tier playlists. There are 46 playlists in this segment. This playlist is excited and contains mostly Indie Rock and Pop music. This playlist just barely touches the 1 million mark for monthly streams. But has the long long stream percentage similar to the high roller segments. Besides the segment for the playlists for personal use, this segemnt also includes the most tracks and the most artists that appear with more than one track in the playlist. This tier contains the playlists that have the highest growth with some playlists having an increase month over month of over 100%




Supervised Analysis
1. Backwards Regression




```{r}

quantile(playlist_data$user_long_stream_ratio, c(.90, .95, .99, 1))

cluster_df$owner_country<- as.factor(cluster_df$owner_country)
cluster_df$genre_1<- as.factor(cluster_df$genre_1)
cluster_df$genre_2<- as.factor(cluster_df$genre_2)
cluster_df$genre_3<- as.factor(cluster_df$genre_3)
cluster_df$mood_1<- as.factor(cluster_df$mood_1)
cluster_df$mood_2<- as.factor(cluster_df$mood_2)
cluster_df$mood_3<- as.factor(cluster_df$mood_3)

size_sample<- floor(0.75 * nrow(cluster_df))

## set the seed to make your partition reproducible
set.seed(123)
train_set <- sample(seq_len(nrow(cluster_df)), size = size_sample)

training <- cluster_df[train_set, ]
testing <- cluster_df[train_set, ]

playlist_full_model<- lm(stream30s ~., data = training[-c(1:3,13,20)])

backwards_model<- stepAIC(playlist_full_model, direction = "backward", 
                      trace = FALSE)

backwards_model$model




```


